Model Search in Relational Domains

### 2016-10-22 experiments.tex
right align numbers
nice symbol names
cleanup experiments

\section{Experimental Results}
We have implemented the instantiation of ACDCL on relational domains
in 2LS verification tool~\cite{2ls}.  We have used the abstract domains
provided by 2LS (intervals, octagons, polyhedra and equalities) for our
experiments. Although the performance of the abstract domain implementations 
in 2LS is not competitive with that of APRON~\cite{apron} library, but 
the domain implementation in 2LS can handle all C operators 
(including bitwise operations) out of the box and support precise
complementation of meet irreducibles, which are necessary for conflict-driven 
learning. Our tool and benchmarks are available online\footnote{http://www.cprover.org/acdcl/}.

We verified a total of 66 benchmarks in ANSI-C, which are derived from 
(a) bit-vector regression category in SV-COMP'16, (c) bit-precise and 
cycle accurate software models of hardware circuits automatically 
generated using v2c~\cite{mtk2016}, (c) controller code generated 
from Simulink (d) several hand-crafted benchmarks for equivalence 
checking and bounded loop analysis.  All bounded loops are completely 
unrolled before analysis.  We compare our tool with a state-of-the-art 
SAT-based bounded model checker CBMC~\cite{cbmc} and a commercial 
static analysis tool, Astr{\'e}e.  CBMC uses MiniSAT solver in the backend.  
Astr{\'e}e is configured with range of abstract domains which includes 
interval, bit-field, congruence, trace partitioning and relational 
domains (octagons, polyhedra, zones, equalities, filter).  Whereas, 
ACDCL is instantiated with a product domain of booleans and template 
polyhedra.  We use an interval and octagon templates for our
experiments.  Additionally, the analysis using ACDCL is configured with 
a heuristic choice of decision (ordered, random, berkmin), 
propagation (forward and multi-way) and conflict-analysis (learning UIP,
DPLL-style).  The timeout for our experiments is set to ~1 hour. 
%
\Omit {
To enable precise analysis using Astr{\'e}e, all our benchmarks are 
manually instrumented with partition directives which provides external 
hint to the tool to guide the trace partitioning heuristics.  Usually, 
such high-precision is not needed for static analysis, since it makes 
the analysis very expensive.  Without trace partitioning, the 
analysis using Astr{\'e}e shows high degree of imprecision. 
}
\paragraph {\textbf{Precision and Efficiency of Analysis}}
Figure~\ref{fig:results} presents a comparison of the analysis 
using CBMC and ACDCL.  Figure~\ref{fig:results}(a) clearly shows that 
the SAT solver made significantly more decisions compared to ACDCL 
for all the benchmarks.  Whereas, the extreme right points below the 
diagonal in Figure~\ref{fig:results}(b) shows the propagations in SAT solver  
is maximum for the benchmarks that exhibit relational behavior.  The 
cumulative statistics corresponding to each phase of the algorithm 
for all 66 benchmarks is presented in Table~\ref{result}.  
Experimental result shows a reduction of at least two orders of 
magnitude in the number of decisions, propagations and conflicts 
compared to analyis using CBMC.    
%
\begin{table}[t]
\begin{center}
{
\begin{tabular}{l|l|l|l|l|l}
\hline
Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
SAT & 27917 & 304031 & 3949 & 33646 & 64 \\ \hline
ACDCL (Product Domain) & 161 & 3130 & 10 & 5 & 0 \\ \hline  
\end{tabular}
}
\end{center}
\caption{Solver statistics}
\label{result}
\end{table}
%
Out of 66 benchmarks, CBMC with MiniSAT backend could only prove 16 benchmarks 
without any restarts.  The solver is restarted in 50 cases to avoid spending 
too much time in branches that do not easily lead to a satisfying assignment or 
stronger clause learning.  In contrast, the analysis using ACDCL solved all
66 benchmarks without any restarts.  This is attributed to decision 
heuristics which exploits the high-level structure of the program combined 
with the stronger deduction and clause learning mechanisms aided by the richer 
abstract domains.  On the other hand, Ast{\'e}e is often faster than ACDCL, 
but the analysis using Astr{\'e}e shows a high degree of imprecision.  
Astr{\'e}e reported 15 false alarms out of 41 safe benchmarks.

\paragraph {\textbf{Propagation Strategy}}      
Figure~\ref{prop-dec}(a) presents a comparison between {\em forward}
propagation strategy and {\em multi-way} propagation strategy in
ACDCL.  The choice of the propagation strategy influences the total
number of decisions and clause learning iterations.  Hence, the
propagation strategy has a significant influence on the run time which
can be visualised in Figure~\ref{prop-dec}(a).  Compared to forward
propagation, the multi-way strategy may take more iterations to reach
the fixed-point, but it significantly reduces the total number of
decisions and conflicts, in particular for the unsafe benchmarks.
This is attributed to the higher precision of the meet irreducibles
inferred by the multi-way strategy which subsequently aids the
decision heuristics to make better decisions.
\pscmt{Chaotic $\rightarrow$ multi-way in Fig.6 }

\paragraph {\textbf{Decision Heuristics}} Figure~\ref{prop-dec}(b) presents a 
visualization of the performance of different decision heuristics in ACDCL.  
Note that the runtimes for all decision heuristics are obtained with multi-way 
propagtion strategy.  The runtimes are very close, but we observe some key
characteristics of these heuristics.  Berkmin heuristic performs consistently 
well for most safe benchmarks and all bit-vector category benchmarks.  Whereas, 
the ordered heuristic performs better for programs with conditional branches 
since it prioritises decisions on meet irreducibles that appear in conditional 
branches over meet irreducibles that involve numerical variables.  Whereas, 
the runtimes for random heuristic is marginally higher than the other two.  This 
suggest that domain-specific decision heuristics are important for ACDCL.

@@

\paragraph {\textbf{Learning}} Learning has a significant influence on
the runtime of ACDCL.  We compare the UIP-based learning technique with 
an analysis that performs chronoligical backtracking without learning. 
The effect of UIP computation allows ACDCL to backtrack non-chronologically 
and guide the model search with a learnt transformer.  However, the DPLL-style 
analysis exhibits case-enumeration behavior and could not finish within the time 
bound for 20\% of our benchmarks.

>>>>

\section{Experimental Results}

We have implemented the instantiation of ACDCL on relational domains in the
2LS verification tool~\cite{2ls}.  We have used the abstract domains
provided by 2LS (intervals, octagons, polyhedra and equalities) for our
experiments.  The performance of the abstract domain implementations in 2LS
is not competitive with that of the APRON~\cite{apron} library; however, the
domain implementation in 2LS handles all C operators (including bitwise
operations) and supports precise complementation of meet irreducibles, which
is necessary for conflict-driven learning.  Our tool and benchmarks are
available online\footnote{http://www.cprover.org/acdcl/} to enable other
researchers to reproduce our results.

We verified a total of 66 ANSI-C benchmarks, which are derived from (a)~the
bit-vector regression category in SV-COMP'16, (b)~bit-precise and
cycle-accurate software models of hardware circuits automatically generated
using v2c~\cite{mtk2016}, (c)~controller code generated from Simulink
(d)~several hand-crafted benchmarks for equivalence checking and bounded
loop analysis.  All bounded loops are completely unrolled before analysis. 
We compare our tool with a state-of-the-art SAT-based bounded model checker
CBMC~\cite{cbmc} and a commercial static analysis tool, Astr{\'e}e. 
\dkcmt{INSERT VERSION NUMBERS OF CBMC AND ASTREE.} CBMC uses MiniSAT in the
backend.  Astr{\'e}e uses a range of abstract domains, which includes
interval, bit-field, congruence, trace partitioning and relational domains
(octagons, polyhedra, zones, equalities, filter).  ACDCL is instantiated
with a product domain of booleans and template polyhedra.  We use the
interval and octagon templates for our experiments.  ACDCL is configured
with a heuristic choice of decision (ordered, random, berkmin), propagation
(forward and multi-way) and conflict-analysis (learning UIP, DPLL-style). 
The timeout for our experiments is set to one hour.
%
\Omit {
To enable precise analysis using Astr{\'e}e, all our benchmarks are 
manually instrumented with partition directives which provides external 
hint to the tool to guide the trace partitioning heuristics.  Usually, 
such high-precision is not needed for static analysis, since it makes 
the analysis very expensive.  Without trace partitioning, the 
analysis using Astr{\'e}e shows high degree of imprecision. 
}

\paragraph {\textbf{Precision and Efficiency of Analysis}}
%
Figure~\ref{fig:results} presents a comparison of the analyses using CBMC
and ACDCL.  Figure~\ref{fig:results}(a) clearly shows that the SAT solver
made significantly more decisions compared to ACDCL for all the benchmarks. 
The points on the extreme right below the diagonal in
Figure~\ref{fig:results}(b) show that the number of propagations in the SAT
solver is maximal for the benchmarks that exhibit relational behavior. 
Cumulative statistics corresponding to each phase of the algorithm for all
66 benchmarks is given in Table~\ref{result}.  We observe a reduction of at
least two orders of magnitude in the number of decisions, propagations and
conflicts compared to analyis using CBMC.

\begin{table}[t]
\begin{center}
{
\begin{tabular}{l|r|r|r|r|r}
\hline
Solver & Decisions & propagations & conflict & conflict literals & restarts \\ \hline
SAT & 27917 & 304031 & 3949 & 33646 & 64 \\ \hline
ACDCL (Product Domain) & 161 & 3130 & 10 & 5 & 0 \\ \hline  
\end{tabular}
}
\end{center}
\caption{Solver statistics}
\label{result}
\end{table}

Out of 66 benchmarks, CBMC with MiniSAT backend could only prove 16
benchmarks without any restarts.  The solver is restarted in 50 cases to
avoid spending too much time in ``hopeless'' branches.  By contrast, the
analysis using ACDCL solved all 66 benchmarks without any restarts.  This is
attributed to our decision heuristics, which exploit the high-level
structure of the program combined with the stronger deduction and clause
learning mechanisms aided by the richer abstract domains.  On the other
hand, Ast{\'e}e is often faster than ACDCL, but the analysis using
Astr{\'e}e shows a high degree of imprecision.  Astr{\'e}e reported 15 false
alarms for the 41 safe benchmarks.

\paragraph {\textbf{Propagation Strategy}}
%
Figure~\ref{prop-dec}(a) presents a comparison between {\em forward}
propagation strategy and {\em multi-way} propagation strategy in ACDCL.  The
choice of the propagation strategy influences the total number of decisions
and clause learning iterations.  Hence, the propagation strategy has a
significant influence on the run time which can be visualised in
Figure~\ref{prop-dec}(a).  Compared to forward propagation, the multi-way
strategy may take more iterations to reach the fixed-point, but it
significantly reduces the total number of decisions and conflicts, in
particular for the unsafe benchmarks.  This is attributed to the higher
precision of the meet irreducibles inferred by the multi-way strategy which
subsequently aids the decision heuristics to make better decisions.

\paragraph {\textbf{Decision Heuristics}}
%
Figure~\ref{prop-dec}(b) presents a visualization of the performance of
different decision heuristics in ACDCL.  Note that the runtimes for all
decision heuristics are obtained with multi-way propagtion strategy.  The
runtimes are very close, but we observe some key characteristics of these
heuristics.  Berkmin heuristic performs consistently well for most safe
benchmarks and all bit-vector category benchmarks.  Whereas, the ordered
heuristic performs better for programs with conditional branches since it
prioritises decisions on meet irreducibles that appear in conditional
branches over meet irreducibles that involve numerical variables.  Whereas,
the runtimes for random heuristic is marginally higher than the other two. 
This suggest that domain-specific decision heuristics are important for
ACDCL.

@@

\paragraph {\textbf{Learning}}
%
Learning has a significant influence on the runtime of ACDCL.  We compare
the UIP-based learning technique with an analysis that performs
chronoligical backtracking without learning.  The effect of UIP computation
allows ACDCL to backtrack non-chronologically and guide the model search
with a learnt transformer.  However, the DPLL-style analysis exhibits
case-enumeration behavior and could not finish within the time bound for
20\% of our benchmarks.
